Figures for Paper

### 2020-11-17 paper_figures.Rmd
tweaks

```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(disease_group2 = case_when(disease == "Control" ~ "Control",
                                 pathology %in% c("FTD-TDP-A","FTD-TDP-B","FTD-TDP-C") ~ "FTLD-TDP",
                                 pathology %in% c("FTD-TAU","FTD-FUS") ~ "FTLD \n non-TDP",
                                 mutations %in% c("SOD1","FUS") ~ "ALS \n non-TDP",
                                 T ~ "ALS-TDP")) %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel 3'", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel 5'", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel 5'")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

@@

boxplot_table %>% 
    filter(disease_tissue == T) %>% 
    group_by(disease_group2) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
    ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion \n reads per million") + 
    theme(text = element_text(size = 18)) + 
    xlab("N samples")

####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "FTLD \n non-TDP",
                              "FTLD-TDP")) %>% 
    filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
    ggplot(aes(x = disease_group2, y = value * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A split \n reads per million") +
    theme(text = element_text(size = 18)) + 
    xlab("")

@@

Splitting by the type of junction in ALS
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "ALS \n non-TDP",
                              "ALS-TDP")) %>% 
    ggplot(aes(x = disease_group2, y = value * 10^6, 
               fill = junction_name,
               group = junction_name)) + 
    geom_boxplot() + 
    geom_point(pch = 21,position = position_dodge(0.8)) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_wrap(vars(tissue_clean)) + 
    ylab("UNC13A cryptic inclusion \n reads per million") +
    # theme(axis.text.x = element_text(angle = 45,hjust=1)) + 
    scale_fill_manual(values = colorblind_pal()(4)[2:4])

@@

new_regress = clean_data_table %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 )

@@

new_regress = new_regress %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf = log10(stmn_2_cryptic_psi_leaf))

>>>>

First we look at detection rate in tissues affected by TDP-43 proteinopathy,
For FTLD this is frontal and temporal Cortices, and
for ALS this is lumbar, cervical, and thoracic spinal cord samples as well as 
motor cortex. For controls we also take all 6 tissues, 
frontal,temporal,motor cortices and the lumbar, cervical, and thoracic spinal cords.

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(disease_group2 = case_when(disease == "Control" ~ "Control",
                                 pathology %in% c("FTD-TDP-A","FTD-TDP-B","FTD-TDP-C") ~ "FTLD-TDP",
                                 pathology %in% c("FTD-TAU","FTD-FUS") ~ "FTLD \n non-TDP",
                                 mutations %in% c("SOD1","FUS") ~ "ALS \n non-TDP",
                                 T ~ "ALS-TDP")) %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel 3'", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel 5'", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel 5'")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

@@

disease_comparisons = list( c("Control","ALS-TDP"),
                           c("Control","ALS \n non-TDP"),
                           c("Control","FTLD-TDP"),
                           c("Control","FTLD \n non-TDP" ))

table_to_test = boxplot_table %>% 
    filter(disease_tissue == T) %>% 
    group_by(disease_group2) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
    dplyr::select(detection_name,
                  participant_id,
                  inclusion_reads,
                  disease_group2) %>% 
    unique()

test_pair = pairwise.wilcox.test(table_to_test$inclusion_reads, table_to_test$detection_name,
                     p.adjust.method = "BH") %>% broom::tidy()

test_pair = test_pair %>% 
  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
                                  p.value < 0.01 ~ "**",
                                   p.value < 0.05 ~ "*",
                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>% 
  mutate(y.position = seq(0.25,by = 0.1,length.out = 9))

table_to_test %>% 
    ggplot(aes(x = detection_name, y = inclusion_reads * 10^6)) + 
    geom_boxplot() + 
    geom_jitter(height = 0) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    ylab("UNC13A cryptic inclusion \n reads per million") + 
    theme(text = element_text(size = 18)) + 
    xlab("N samples") + 
    stat_pvalue_manual(test_pair %>% filter(p.value < 0.05),
                       label = "p_value_draw") +
   stat_compare_means()

####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "FTLD \n non-TDP",
                              "FTLD-TDP")) %>% 
    filter(tissue_clean %in% c("Cerebellum","Frontal Cortex","Temporal Cortex")) %>% 
  dplyr::select(disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
  unique() %>% 
    ggplot(aes(x = disease_group2, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F) + 
    geom_jitter(height = 0,show.legend = F) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 18)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4])

@@

Splitting by the type of junction in ALS
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table %>% 
  filter(disease_group2 %in% c("Control",
                              "ALS \n non-TDP",
                              "ALS-TDP")) %>% 
      filter(tissue_clean %in% c("Cerebellum","Motor Cortex","Temporal Cortex") | grepl("Cord",tissue_clean)) %>% 
  dplyr::select(disease_group2,
                tissue_clean,
                junction_name,
                value) %>% 
  unique() %>% 
  mutate(disease_group2 = fct_relevel(disease_group2,"Control","ALS \n non-TDP","ALS-TDP")) %>% 
    ggplot(aes(x = disease_group2, 
               y = value * 10^6,
               fill = junction_name)) + 
    geom_boxplot(show.legend = F) + 
    geom_jitter(height = 0,show.legend = F) + 
    scale_y_log10() + 
    ggpubr::theme_pubr() +
    facet_grid(vars(junction_name),vars(tissue_clean)) + 
    ylab("UNC13A cryptic \n reads per million") +
    theme(text = element_text(size = 18)) + 
    xlab("") +
    scale_fill_manual(values = colorblind_pal()(4)[2:4]) +
  theme(axis.text.x =  element_text(size = 14))
```

@@

new_regress = clean_data_table %>% 
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>%
    filter(STMN2_annotated_leaf > 10) %>% 
    filter(stmn_2_cryptic_psi_leaf > 0 ) %>%  
    mutate(log10_stmn_2_cryptic_psi_leaf = log10(stmn_2_cryptic_psi_leaf)) %>% 
    mutate(log10_unc13a_cryptic_leaf_psi_plusone = log10(unc13a_cryptic_leaf_psi + 1)) %>% 
    mutate(log10_stmn_2_cryptic_psi_leaf_plusone = log10(stmn_2_cryptic_psi_leaf + 1))

@@
